Equivalence Testing and the Second Generation P-Value

### 2018-08-17 sgpv_vs_TOST.Rmd
new version feedback Marie

To test predictions most researchers predominantly rely on null-hypothesis tests. 
This statistical approach can be used to examine whether observed data is sufficiently surprising under the null hypothesis to reject an effect size of zero. 
Null-hypothesis tests have an important limitation, in that this procedure can only reject the hypothesis that there is no effect, while scientists should also be able to provide statistical support for *equivalence*. 
When testing for equivalence researchers aim to examine whether the effect is practically zero, or that there is no meaninful effect. 
By specifying a range around the the null hypothesis of values that are deemed practically equivalent to the absence of an effect (i.e., 0 ± 0.3) the observed data can be compared against an *equivalence range* and researchers can test if a meaningful effect is absent (e.g., Hauk & Anderson, 1985; Kruschke, 2018; Metzler, 1974; Rogers, Howard, & Vessey, 1993; Serlin & Lapsey, 1985; Spiegelhalter, Freedman, & Parmar, 1994; Wellek, 2010; Westlake, 1972).

@@

#The relationship between *p*-values from TOST and SGPV

To examine the relation between the TOST p-value and the SGPV we can calculate both statistics across a range of observed effect sizes. In Figure 1 *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for one-sample *t*-tests where the observed mean in the sample ranges from 140 to 150 (on the x-axis), the equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), a standard deviation of 2 and a sample size of 100.  Our conclusions about the relationship between TOST *p*-values and SGPV in this article are not dependent upon any specific example.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "",
     xlab = "Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

lines(sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
#abline(h = c(0.975, 0.5, 0.025), col = "grey", lwd = 1, lty = 3)
```
*Figure 1*: Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 100 and a standard deviation of 2.

The SGPV treats the equivalence range as the null-hypothesis, while the TOST procedure treats the values outside of the equivalence range as the null-hypothesis. For ease of comparison we can reverse the SGPV (by calculating 1-SGPV) to make the values more easily comparable. We see that the *p*-value from the TOST procedure and the SGPV follow each other closely. 
```{r 1-sgpv_tost, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "SGPV or TOST p-value",
     xlab = "Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

lines(1-sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
#abline(h = c(0.975, 0.5, 0.025), col = "grey", lwd = 1, lty = 3)
```
```{r, include=FALSE}
m <- 145
mu <- 145
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

m <- 140
mu <- 145
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
```
*Figure 2*: Comparison of *p*-values from TOST (black line) and 1-SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 100 and a standard deviation of 2.

When the observed sample mean is 145 and we are testing against equivalence bounds of 143 and 147 using the TOST procedure for a one-sample *t*-test with a sample size of 100 and a standard deviation of 2, the equivalence test is significant, *t*(`r tost_res1$TOST_df`) = `r tost_res1$TOST_t1`, *p* = `r tost_res1$TOST_p1`. Because the 95% CI falls completely within the equivalence bounds, the SGPV is `r SGPV1` (see Figure 1).

One the other hand, if the observed mean is 140, the equivalence test is not significant (the observed mean is far outside the equivalence range of 143 to 147), *t*(`r tost_res2$TOST_df`) = `r tost_res2$TOST_t1`, *p* = `r tost_res2$TOST_p1` (or more accurately, *p* > .999 as *p*-values are bounded between 0 and 1). Because the 95% CI falls completely outside the equivalence bounds, the SGPV is `r SGPV2` (see Figure 1).

@@

It is clear the SGPV and the *p*-value from TOST are closely related. We can think of the SGPV as a straight line that will always overlap the *p*-value from an equivalence test in 3 points. When the TOST *p*-value is 0.5, the SGPV is also 0.5 (note that the reverse is not true). The SGPV is 50% when the observed mean falls exactly on the lower or upper equivalence bound. When the observed mean equals the equivalence bound, the difference between the mean in the data and the equivalence bound is 0, the *t*-value for the equivalence test is also 0, and thus the *p*-value is 0.5 (situation A). 
```{r, include=FALSE}
m <- 146.5
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.025

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
SGPV2

@@

tost_res3 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV3 <- TOST_to_SGPV(tost_res = tost_res3)
SGPV3

@@

Two other points always have to overlap. When the 95% CI falls completely, but only just inside the equivalence region, the TOST (which relies on a one-sided test) should be significant at an alpha level of 0.025. When the SGPV changes from <1 to 1 the 95% CI is exactly equal to one of the equivalence bounds (see situation B in the plot above, where the 95% CI falls completely inside the equivalence bounds) the TOST *p*-value is 0.025. The third point where the SGPV and the *p*-value from the TOST procedure should overlap is where the SGPV  changes from a positive value (i.e., 0.0001) to 0 (when the 95% CI completely falls outside of the equivalence bound, see situation C in the plot above). When the 95% CI touches the outside of the equivalence bound and the TOST *p*-value will be 0.975.

The confidence interval width is a uniformly distributed across the mean differences, in the sense that as the true mean in a one-sample t-test gets closer to the test value (in the plot below, from situation A to D, the mean gets closer to the test value by 0.1) the difference in the overlap is stable.
```{r, include=FALSE}
m <- 146
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)

@@

tost_res3 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV3 <- TOST_to_SGPV(tost_res = tost_res3)

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
```{r, echo=FALSE, dpi=600, fig.width=5, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(-3, 3),
     yaxt = "n",
     ylab = "",
     xlab = "Mean Difference")
axis(2, at = c(0.2,0.4,0.6,0.8), labels = c("D", "C", "B", "A"), las = 1)
abline(v = tost_res1$high_eqbound, 
       lty = 2)
abline(v = tost_res1$low_eqbound, 
       lty = 2)
abline(v = 0, 
       lty = 2, 
       col = "grey")

@@

par(new=TRUE)
curve(dnorm(x, 1.2, 0.5), from=-0.3, to=2.7, ylim = c(-2, 8), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.3, 0.5), from=-0.2, to=2.8, ylim = c(-4, 6), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.4, 0.5), from=-0.1, to=2.9, ylim = c(-6, 4), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.5, 0.5), from=0, to=3, ylim = c(-8, 2), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
```
*Figure 4*: Means, normal distribution, and 95% CI for data with a sample size of 1000000 and a standard deviation of 500 for samples where the true population mean is 1.5, 1.4, 1.3, and 1.2.

@@

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
For example, the SGPV from A to D is `r SGPV1`, `r SGPV2`, `r SGPV3`, and `r SGPV4`. The difference in the percentage of overlap between A and B (`r SGPV1-SGPV2`) is identical to the difference in the percentage of overlap between C and D as the mean gets 0.1 closer to the test value (`r SGPV3-SGPV4`). 
```{r , echo=FALSE}
tail1 <- 1-pnorm(2, 1.5, 0.5)
tail2 <- 1-pnorm(2, 1.4, 0.5)
tail3 <- 1-pnorm(2, 1.3, 0.5)
tail4 <- 1-pnorm(2, 1.2, 0.5)
```

As we move the means closer to the test value in steps of 0.1 across A to D the *p*-value calculated for normally distributed data is not uniformly distributed. The probability of observing data more extreme than the upper bound of 2 is (from A to D) `r tail1`, `r tail2`, `r tail3`, and `r tail4`. As we can see, the difference between A and B (`r tail1-tail2`) is not the same as the difference between C And D (`r tail3-tail4`). Indeed, the difference in *p*-values is the largest as you start at *p* = 0.5 (when the observed mean falls on the test value), which is why the line in Figure 1 is the steepest at *p* = 0.5. Note that where the SGPV reaches 1 or 0, *p*-values closely approximate 0 and 1, but never reach these values.

@@

  invisible(capture.output(res <- TOSTone.raw(m = m, 
                                              mu = mu,
                                              sd = sd, 
                                              n = n, 
                                              low_eqbound = low_eqbound, 
                                              high_eqbound = high_eqbound, 
                                              alpha = alpha,
                                              plot = FALSE
  )))
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(sgpv_list, 
     p_tost_list,
     type="l",
     lwd = 3, 
     ylim = c(0, 1), 
     xlim = c(0, 1),
     # yaxt = "n",
     # xaxt = "n",
     ylab = "TOST p-value",
     xlab = "SGPV")
```
*Figure 5*: The relationship between *p*-values from the TOST procedure and the SGPV for the same scenario as in Figure 1.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 600),
     yaxt = "n",
     xaxt = "n",
     ylab = "",
     xlab = "Mean")
axis(1, at = seq(0,600,100), labels = seq(-3,3,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

@@

We can again plot the two statistics against each other so see where they are unrelated (indicated by straight lines in the curve). We see the SGPV is 0.5 for a range of observed means where the p-value from the equivalence test still varies. It should be noted that in these calculations the *p*-values for the TOST procedure are *never* smaller than 0.05 (i.e., they do not get below 0.05 on the y-axis). In other words, we can't conclude equivalence based on any of the observed means. How is this possible? Remember that the TOST procedure consists of two one-sided tests against the upper and lower equivalence bound. The TOST *p*-value is smaller than 0.05 if the 90% CI falls completely between the upper and lower equivalence bounds. However, we are examining a scenario where the 90% CI is so wide that it never falls completely within the two equivalence bounds. As Lakens (2017) notes: “in small samples (where CIs are wide), a study might have no statistical power (i.e., the CI will always be so wide that it is necessarily wider than the equivalence bounds).” None of the *p*-values based on the TOST procedure are below 0.05, and thus, in the long run we have 0% power.
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(sgpv_list, 
     p_tost_list,
     type="l",
     lwd = 3, 
     ylim = c(0, 1), 
     xlim = c(0, 1),
     # yaxt = "n",
     # xaxt = "n",
     ylab = "TOST p-value",
     xlab = "SGPV")
```
*Figure 7*: The relationship between *p*-values from the TOST procedure and the SGPV for the same scenario as in Figure 6.

>>>>

To test predictions most researchers predominantly rely on null-hypothesis tests. 
This statistical approach can be used to examine whether observed data is sufficiently surprising under the null hypothesis to reject an effect size of zero. 
Null-hypothesis tests have an important limitation, in that this procedure can only reject the hypothesis that there is no effect, while scientists should also be able to provide statistical support for *equivalence*. 
When testing for equivalence researchers aim to examine whether the effect is practically zero, or that there is no meaninful effect. 
By specifying a range around the null hypothesis of values that are deemed practically equivalent to the absence of an effect (i.e., 0 ± 0.3) the observed data can be compared against an *equivalence range* and researchers can test if a meaningful effect is absent (e.g., Hauk & Anderson, 1985; Kruschke, 2018; Metzler, 1974; Rogers, Howard, & Vessey, 1993; Serlin & Lapsey, 1985; Spiegelhalter, Freedman, & Parmar, 1994; Wellek, 2010; Westlake, 1972).

@@

#The relationship between *p*-values from TOST and SGPV when confidence intervals are symmetrical

To examine the relation between the TOST p-value and the SGPV we can calculate both statistics across a range of observed effect sizes. In Figure 1 *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for hypothetical one-sample *t*-tests for *all* means that can be observed in studies ranging from 140 to 150 (on the x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), the observed standard deviation is assumed to be 2, and the sample size is 100. For example, for the left-most point in Figure 1 the SGPV and the TOST *p*-value is calculated for a hypothetical study with a sample size of 100, an observed standard deviation of 2, and an observed mean of 140, where the *p*-value for the equivalence test is 1, and the SGPV is 0. Our conclusions about the relationship between TOST *p*-values and SGPV in this article are not dependent upon any specific example.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

lines(sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
#abline(h = c(0.975, 0.5, 0.025), col = "grey", lwd = 1, lty = 3)
```
*Figure 1*: Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2.

The SGPV treats the equivalence range as the null-hypothesis, while the TOST procedure treats the values outside of the equivalence range as the null-hypothesis. For ease of comparison we can reverse the SGPV (by calculating 1-SGPV) to make the values more easily comparable. We see that the *p*-value from the TOST procedure and the SGPV follow each other closely. 
```{r 1-sgpv_tost, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and 1-SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

lines(1-sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
#abline(h = c(0.975, 0.5, 0.025), col = "grey", lwd = 1, lty = 3)
```
```{r, include=FALSE}
m <- 145
mu <- 145
sd <- 2
n <- 30
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

m <- 140
mu <- 145
sd <- 2
n <- 30
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
```
*Figure 2*: Comparison of *p*-values from TOST (black line) and 1-SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2.

When the observed sample mean is 145 and we are testing against equivalence bounds of 143 and 147 using the TOST procedure for a one-sample *t*-test with a sample size of 30 and a standard deviation of 2, the equivalence test is significant, *t*(`r tost_res1$TOST_df`) = `r round(tost_res1$TOST_t1, digits = 2)`, *p* = `r round(tost_res1$TOST_p1, digits =  4)`. Because the 95% CI falls completely within the equivalence bounds, the SGPV is `r SGPV1` (see Figure 1).

On the other hand, if the observed mean is 140, the equivalence test is not significant (the observed mean is far outside the equivalence range of 143 to 147), *t*(`r tost_res2$TOST_df`) = `r round(tost_res2$TOST_t1, digits = 2)`, *p* = `r round(tost_res2$TOST_p1, digits =  4)` (or more accurately, *p* > .999 as *p*-values are bounded between 0 and 1). Because the 95% CI falls completely outside the equivalence bounds, the SGPV is `r SGPV2` (see Figure 1).

@@

It is clear the SGPV and the *p*-value from TOST are closely related. We can think of the SGPV as a straight line that will always overlap the *p*-value from an equivalence test in 3 points. When the TOST *p*-value is 0.5, the SGPV is also 0.5 (note that the reverse is not true). The SGPV is 50% when the observed mean falls exactly on the lower or upper equivalence bound. When the observed mean equals the equivalence bound, the difference between the mean in the data and the equivalence bound is 0, the *t*-value for the equivalence test is also 0, and thus the *p*-value is 0.5 (situation A, Figure 3). 
```{r, include=FALSE}
m <- 146.5
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.025

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)
SGPV2

@@

tost_res3 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV3 <- TOST_to_SGPV(tost_res = tost_res3)
SGPV3

@@

Two other points always have to overlap. When the 95% CI falls completely, but only just inside the equivalence region, the TOST (which relies on a one-sided test) should be significant at an alpha level of 0.025. When the SGPV changes from <1 to 1 the 95% CI is exactly equal to one of the equivalence bounds (see situation B in Figure 3, where the 95% CI falls completely inside the equivalence bounds) the TOST *p*-value is 0.025. The third point where the SGPV and the *p*-value from the TOST procedure should overlap is where the SGPV  changes from a positive value (i.e., 0.0001) to 0 (when the 95% CI completely falls outside of the equivalence bound, see situation C in Figure 3). When the 95% CI touches the outside of the equivalence bound and the TOST *p*-value will be 0.975.

The confidence interval width is uniformly distributed across the mean differences, in the sense that as the true mean in a one-sample t-test gets closer to the test value (in the plot below, from situation A to D, the mean gets closer to the test value by 0.1) the difference in the overlap is stable.
```{r, include=FALSE}
m <- 146
mu <- 144.5
sd <- 500
n <- 1000000
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

@@

tost_res2 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV2 <- TOST_to_SGPV(tost_res = tost_res2)

@@

tost_res3 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV3 <- TOST_to_SGPV(tost_res = tost_res3)

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
```{r, echo=FALSE, dpi=600, fig.width=5, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(-3, 3),
     yaxt = "n",
     ylab = "",
     xlab = "Mean Difference")
axis(2, at = c(0.2,0.4,0.6,0.8), labels = c("D", "C", "B", "A"), las = 1)
abline(v = tost_res1$high_eqbound, 
       lty = 2)
abline(v = tost_res1$low_eqbound, 
       lty = 2)
abline(v = 0, 
       lty = 2, 
       col = "grey")

@@

par(new=TRUE)
curve(dnorm(x, 1.2, 0.5), from=-0.3, to=2.7, ylim = c(-2, 8), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.3, 0.5), from=-0.2, to=2.8, ylim = c(-4, 6), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.4, 0.5), from=-0.1, to=2.9, ylim = c(-6, 4), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
par(new=TRUE)
curve(dnorm(x, 1.5, 0.5), from=0, to=3, ylim = c(-8, 2), xlim = c(-3, 3), xlab = "", ylab = "", axes = FALSE)
```
*Figure 4*: Means, normal distribution, and 95% CI for samples where the true population mean is 1.5, 1.4, 1.3, and 1.2.

@@

tost_res1 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV1 <- TOST_to_SGPV(tost_res = tost_res1)

@@

tost_res4 <- TOSTone.raw(m = m, 
                   mu = mu,
                   sd = sd, 
                   n = n, 
                   low_eqbound = low_eqbound, 
                   high_eqbound = high_eqbound, 
                   alpha = alpha,
                   plot = FALSE,
                   verbose = FALSE
)
SGPV4 <- TOST_to_SGPV(tost_res = tost_res4)
```
For example, the SGPV from A to D is `r round(SGPV1, digits = 2)`, `r round(SGPV2, digits = 2)`, `r round(SGPV3, digits = 2)`, and `r round(SGPV4, digits = 2)`. The difference in the percentage of overlap between A and B (`r round(SGPV1-SGPV2, digits = 2)`) is identical to the difference in the percentage of overlap between C and D as the mean gets 0.1 closer to the test value (`r round(SGPV3-SGPV4, digits = 2)`). 
```{r , echo=FALSE}
tail1 <- 1-pnorm(2, 1.5, 0.5)
tail2 <- 1-pnorm(2, 1.4, 0.5)
tail3 <- 1-pnorm(2, 1.3, 0.5)
tail4 <- 1-pnorm(2, 1.2, 0.5)
```

As we move the means closer to the test value in steps of 0.1 across A to D the *p*-value calculated for normally distributed data is not uniformly distributed. The probability of observing data more extreme than the upper bound of 2 is (from A to D) `r round(tail1, digits = 3)`, `r round(tail2, digits = 3)`, `r round(tail3, digits = 3)`, and `r round(tail4, digits = 3)`. As we can see, the difference between A and B (`r round(tail1-tail2, digits = 3)`) is not the same as the difference between C And D (`r round(tail3-tail4, digits = 3)`). Indeed, the difference in *p*-values is the largest as you start at *p* = 0.5 (when the observed mean falls on the test value), which is why the line in Figure 1 is the steepest at *p* = 0.5. Note that where the SGPV reaches 1 or 0, *p*-values closely approximate 0 and 1, but never reach these values.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(sgpv_list, 
     p_tost_list,
     type="l",
     lwd = 3, 
     ylim = c(0, 1), 
     xlim = c(0, 1),
     # yaxt = "n",
     # xaxt = "n",
     ylab = "TOST p-value",
     xlab = "SGPV")
```
*Figure 5*: The relationship between *p*-values from the TOST procedure and the SGPV for the same scenario as in Figure 1.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 600),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,600,100), labels = seq(-3,3,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

@@

We can again plot the two statistics against each other to see where they are unrelated (indicated by straight lines in the curve). We see the SGPV is 0.5 for a range of observed means where the p-value from the equivalence test still varies. It should be noted that in these calculations the *p*-values for the TOST procedure are *never* smaller than 0.05 (i.e., they do not get below 0.05 on the y-axis). In other words, we cannot conclude equivalence based on any of the observed means. How is this possible? Remember that the TOST procedure consists of two one-sided tests against the upper and lower equivalence bound. The TOST *p*-value is smaller than 0.05 if the 90% CI falls completely between the upper and lower equivalence bounds. However, we are examining a scenario where the 90% CI is so wide that it never falls completely within the two equivalence bounds. As Lakens (2017) notes: “in small samples (where CIs are wide), a study might have no statistical power (i.e., the CI will always be so wide that it is necessarily wider than the equivalence bounds).” None of the *p*-values based on the TOST procedure are below 0.05, and thus, in the long run we have 0% power.
```{r, echo=FALSE, dpi=600, fig.width=6, fig.height=5}
plot(sgpv_list, 
     p_tost_list,
     type="l",
     lwd = 3, 
     ylim = c(0, 1), 
     xlim = c(0, 1),
     # yaxt = "n",
     # xaxt = "n",
     ylab = "TOST p-value",
     xlab = "SGPV")
```
*Figure 7*: The relationship between *p*-values from the TOST procedure and the SGPV for the same scenario as in Figure 6.
